---
title: "data_wrangling_i"
output: github_document
---
```{r}
library(tidyverse)
```

```{r}
library(here)
here("data", "file_i_want.csv")
```

'getwd()' means get working directory.
"setwd()' means get absolute directory.

| Symbol | Meaning | Example from this working directory |
|---|---|---|
| `~` | your home directory | `/Users/yan` |
| `.` | current working directory | `/Users/yan/Desktop/p8105_project` |
| `..` | one folder up | `/Users/yan/Desktop` |
| `../..` | two folders up | `/Users/yan` |

```{r}
litters_df = read_csv(file = "./data_import_examples/FAS_litters.csv")
```
```{r}
names(litters_df)
litters_df = janitor::clean_names(litters_df)
# janitor::clean_names can help to unify the format column names
names(litters_df)
```
```{r}
library(conflicted)
filter(group == "con7")
```

can help to remind the conflict in r package, 
eg: you use 'filter' during the code, then it will alert:

stats::filter / dplyr::filter

# Looking at data
```{r}
view(litters_df)
str(litters_df)
head(litters_df,5)
tail(litters_df,5)
litters_df
```

More neat than useful
```{r}
skimr::skim(litters_df)
```

If your file is not standard or it is messy, you can use 'read_*' to read the file with parameters setting to tell R how to read.

## Col_names:

col_names = TRUE means treating first raw as column names.

col_names = FALSE means treating first raw as regular data, then it will generate column name, like x1, x2,...  for the data set.

# skip:
number of rows to skip before reading data.

```{r}
litters_df = 
    read_csv(file = "./data_import_examples/FAS_litters.csv",
    skip = 10, col_names = FALSE)
head(litters_df)
```

## na: 

It tells r which symbol represent missing value
eg: "."," ","na","NA","0",etc.

```{r}
litters_df = 
    read_csv(
        file = "./data_import_examples/FAS_litters.csv",
        na = c(".", "NA", ""))
head(litters_df)
```

```{r}
litters_df = 
    read_csv(file = "./data_import_examples/FAS_litters.csv",
        na = c(".", "NA", ""),
    col_types = cols(
      Group = col_character(),
      `Litter Number` = col_character(),
      `GD0 weight` = col_double(),
      `GD18 weight` = col_double(),
      `GD of Birth` = col_integer(),
      `Pups born alive` = col_integer(),
      `Pups dead @ birth` = col_integer(),
      `Pups survive` = col_integer()
    )
  )

tail(litters_df)
```
Note that you don’t have to specify the variable type for every column, and can only focus on ones that are difficult:

```{r}
litters_df = 
    read_csv(file = "./data_import_examples/FAS_litters.csv",
        na = c(".", "NA", ""),
    col_types = cols(
      Group = col_factor()
    )
)

head(litters_df)
```

# Learning Assessment:
```{r}
pups_df = read_csv(file = "./data_import_examples/Fas_pups.csv",
          skip = 3, 
          na = c(".","NA"),
          col_types = "fdddddd")
# names(pups_df)
skimr::skim(pups_df)
```
# import excel

```{r}
library(readxl)

mlb11_df = read_excel("./data_import_examples/mlb11.xlsx", n_max = 20)

head(mlb11_df, 5)
```

# import SAS

```{r}
library(haven)

pulse_df = read_sas("./data_import_examples/public_pulse_data.sas7bdat")

head(pulse_df, 5)
```

Learning Assessment: 
```{r}
library(readr)
pups_base = read_csv("./data_import_examples/FAS_pups.csv", skip = 3)
pups_readr = read.csv("./data_import_examples/FAS_pups.csv", skip = 3)
# read_csv() numeric columns → often shown as <dbl>
# read.csv() numeric columns → may be stored as <int> if all values are integers
View(pups_base)
View(pups_readr)

pups_base
pups_readr

pups_base$S
pups_readr$S
```
Why pups_readr$S can return Sex column:

Because tibbles created by read_csv() allow partial matching behavior in $ lookup in this context.

Base R 'read.csv()' automatically converts non-syntactic names by default:
Litter Number → Litter.Number
PD ears       → PD.ears

whereas 'read_csv()' preserves the original names.
Litter Number
PD ears

*prefer read_csv*

